---
title: Simple Linear Regression Analysis for Prediction of the % of Student based
  on the Number of Study Hour
author: "Monone Damilola Elizabeth"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(ggplot2)
```

## Introduction
This document performs a simple linear regression analysis to predict student scores based on study hours.

# Data Loading
```{r Load the dataset}
Hours_Scores <- read_excel("Hours_Scores.xlsx")

# Print the first few rows of the data
head(Hours_Scores)
```

## Exploratory Data Analysis (EDA)
To check the structure and summary statistics of the data.

```{r Check the column names}
# View the first few rows of the data
head(Hours_Scores)

# Summary statistics
summary(Hours_Scores)

# Structure of the data
str(Hours_Scores)
```

# Visualization of the data
```{r Scatter plot}
ggplot(Hours_Scores, aes(x=Hours, y=Scores)) + 
  geom_point() + 
  labs(title="Hours vs Scores", x="Study Hours", y="Percentage Score")
```

The plottted scatter plot of the data is to visualize the relationship between study hours and scores. 

## Linear Regression Model
```{r Fit the linear model}
model <- lm(Scores ~ Hours, data=Hours_Scores)

# Summary of the model
print(summary(model))

# Plot the regression line
ggplot(Hours_Scores, aes(x=Hours, y=Scores)) + 
  geom_point() + 
  geom_smooth(method="lm", col="blue") + 
  labs(title="Hours vs Scores with Regression Line", x="Study Hours", y="Percentage Score")
```

The plotted regression line and the scatter points shows a possitive realtionship between the study hours and the percentage scores of the students.

## Predictions and Residuals
```{r Make predictions}
predicted_scores <- predict(model, newdata=Hours_Scores)

# Add predictions to the data frame
Hours_Scores$Predicted_Scores <- predicted_scores

# Calculate residuals
Hours_Scores$Residuals <- Hours_Scores$Scores - Hours_Scores$Predicted_Scores

# Print the first few rows to verify predictions and residuals
print(head(Hours_Scores))
```

## Observed vs Predicted Plot
```{r Plot the observed vs predicted values}
ggplot(Hours_Scores, aes(x=Hours, y=Scores)) + 
  geom_point() + 
  geom_line(aes(y=Predicted_Scores), color="blue") + 
  labs(title="Observed vs Predicted Scores", x="Study Hours", y="Scores")
```

## Model Evaluation
```{r To calculate the RMSE and MAE}
# Root Mean Squared Error (RMSE)
rmse <- sqrt(mean(Hours_Scores$Residuals^2))
print(paste("RMSE:", rmse))

# Mean Absolute Error (MAE)
mae <- mean(abs(Hours_Scores$Residuals))
print(paste("MAE:", mae))
```
## Predicting New Data
```{r Prediction score for a student who studies 9.25 hours per day}
new_data <- data.frame(Hours = 9.25)
predicted_score <- predict(model, newdata=new_data)
print(paste("Predicted score for a student who studies 9.25 hours per day:", predicted_score))
```
## Conclusion
I performed a simple linear regression analysis to predict student scores based on study hours. I evaluated the model using RMSE and MAE, and made predictions for new data. 
From the analysis, we can draw the following conclusions:
1. The analysis confirms that increasing the number of study hours has a positive impact on student scores. This relationship is statistically significant and can be used to predict scores based on study hours.
2. The linear regression model fits the data reasonably well, as indicated by the R-squared value and the residual standard error. However, it is important to consider other potential variables that might influence student scores, which are not included in this simple model.
3. The RMSE and MAE values provide a quantitative measure of the model's prediction accuracy. While these metrics indicate that the model performs reasonably well, there is still some room for improvement, as reflected by the residuals.
4. The model can be used to make informed predictions about student scores based on the number of study hours. For example, a student studying 9.25 hours per day is predicted to achieve a specific score, which can help in planning study schedules and setting academic goals.

